---
title: 'statistical inference : assignment 2'
author: "vincent trouillet"
date: "7 mars 2018"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

## Overview

The purpose of the this data analysis is to analyze the ToothGrowth data set by comparing the guinea tooth growth by supplement and dose. A rapid exploratory data analysis on the data set and a comparison with confidence intervals  will be made in order to make conclusions about the tooth growth.

## 1) The data : basic exploratory data analyses

```{r data, echo=TRUE}
library(datasets)
library(ggplot2)

data(ToothGrowth)

head(ToothGrowth)
str(ToothGrowth)
summary(ToothGrowth)


```

Total Observations are 60 with 3 columns - len, supp, dose

Tooth Length (len) is of type numeric values.

Supplementary (supp) is of type OJ(orange juice) and VC(vitamin C)

Dosage (dose) is of 3 Types - 0.5, 1 and 2

## 3) Comparison of tooth growth by supp and dose


### use boxplot for a graphical comparison

```{r plot, echo=TRUE}

levels(ToothGrowth$supp)<-c("Orange Juice","Vitamin C")
plot.bxpl <- ggplot(ToothGrowth, aes(x=dose,y=len,fill=factor(dose))) +
    facet_grid(.~supp)+ geom_boxplot() + 
    scale_x_discrete("Dose (mg/day)") +   
    scale_y_continuous("Tooth Length") +  
    ggtitle("Tooth Length by Dosage for each type of supplement")

plot.bxpl


```

We  can see that there is a relation between dose and tooth growth. Orange Juice seems more effective for dosages 0.5 & 1.0. For a dose amount 2.0 mg/day, both supplement look equivalent. 

let's explore hypothesis.

We assume that no other unmeasured factors are affecting tooth length and the tooth lengths follow a normal distrubution.

### Hypothese 1 : orange juice & vitamin C deliver the same tooth growth across the dataset

```{r hyp1, echo=TRUE}

hypoth1 <- t.test(len ~ supp, data = ToothGrowth)
hypoth1
```

The confidence intervals includes 0 and the p-value is greater than the threshold of 0.05. The null hypothesis cannot be rejected.

### Hypothese 2 : orange juice & vitamin C deliver the same tooth growth for a dosage of 0.5 mg/day

```{r hyp2, echo=TRUE}
hypoth2<-t.test(len ~ supp, data = subset(ToothGrowth, dose == 0.5))
hypoth2
```

The confidence interval does not include 0 and the p-value is below the 0.05 threshold. The null hypothesis can be rejected. The alternative hypothesis that 0.5 mg/day dosage of orange juice delivers more tooth growth than ascorbic acid is accepted.

### Hypothese 3 : orange juice & vitamin C deliver the same tooth growth for a dosage of 1 mg/day

```{r hyp3, echo=TRUE}

hypoth3<-t.test(len ~ supp, data = subset(ToothGrowth, dose == 1))
```

The confidence interval does not include 0 and the p-value is smaller than the 0.05 threshold. The null hypothesis can be rejected. The alternative hypothesis that 1 mg/day dosage of orange juice delivers more tooth growth than ascorbic acid is accepted.

### Hypothese 4 : orange juice & vitamin C deliver the same tooth growth for a dosage of 2 mg/day

```{r hyp4, echo=TRUE}
hypoth4<-t.test(len ~ supp, data = subset(ToothGrowth, dose == 2))
hypoth4
```

The confidence interval does include 0 and the p-value is larger than the 0.05 threshold. The null hypothesis cannot be rejected.

## 4) Conclusions

Orange juice delivers more tooth growth than ascorbic acid for dosages 0.5 & 1.0. Orange juice and ascorbic acid deliver the same amount of tooth growth for dose amount 2.0 mg/day. For the entire data set we cannot conclude orange juice is more effective that ascorbic acid.


